---
title: "Homework #1"
author: "David Nemirovsky"
output: github_document
date: 2/14/21
--- 

```{r setup, include=FALSE}
library(tidyverse)
library(ISLR)
library(glmnet)
library(caret)
library(corrplot)
library(plotmo)

knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(37564)
```

In this exercise, we will predict solubility of compounds using their chemical structures. The training data are in the file "solubility_train.csv" and the test data are in "solubility_test.csv". Among the 228 predictors, 208 are binary variables that indicate the presence or absence of a particular chemical substructure, 16 are count features, such as the number of bonds or the number of bromine atoms, and 4 are continuous features, such as molecular weight or surface area. The response is in the column "Solubility" (the last column).

 (a) Fit a linear model using least squares on the training data and calculate the mean squared error using the test data.

+ First, let's read in the data:

```{r read, message = F}
train_df = read_csv("solubility_train.csv")
test_df = read_csv("solubility_test.csv")
```

+ Next, let's create a linear model with 10-fold CV:

```{r lm, warning = F}
set.seed(37564)
ctrl = trainControl(method = "cv", number = 10)
model_lm = train(Solubility~., 
               data = train_df, 
               method = "lm", 
               trControl = ctrl)

model_lm$finalModel$coefficients %>% 
  broom::tidy() %>% 
  knitr::kable(col.names = c("Predictor", "Beta Value"))
```

+ Now, let's calculate the MSE using the test data:

```{r mse lm}
lm_pred = predict(model_lm, newdata = test_df)
mse_lm = mean((lm_pred - test_df$Solubility)^2)
```

+ Therefore, the MSE of this linear model is **`r round(mse_lm, 4)`**.